Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-01-21 ms/data_readme.Rmd
Change "dryad" to "figshare"

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ebihara_2019_zip_file",
    "ftol_zip_file",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_dryad_file",
    "comm_ferns_dryad_file",
    "shape_ferns_dryad_file",
    "fern_traits_dryad_file",
    "japan_fern_phylogram_dryad_file",
    "japan_fern_tree_dryad_file",
    "japan_fern_tree_uncollapsed_dryad_file",
    "refs_yaml"
    )
  )
```

@@

```{r japan-ferns-biodiv-dryad}
japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
```

`r fs::path_file(japan_ferns_biodiv_dryad_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. 
Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Includes one row with missing
environmental data and one outlier for % apomixis that were removed prior to
spatial modeling analysis in Nitta et al. 2021.

Number of variables: `r ncol(japan_ferns_biodiv_dryad)`

Number of cases/rows: `r nrow(japan_ferns_biodiv_dryad)`

@@

```{r japan-ferns-biodiv-dryad-cols, results = "asis"}
  japan_ferns_biodiv_dryad %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "lat", "Latitude of grid-cell centroid",
      "long", "Longitude of grid-cell centroid",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "fd_obs_z", "Z-score (standard effect size) of observed functional diversity (FD)",
      "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
      "pe_obs_z", "Z-score of observed phylogenetic endemisim (PE)",
      "rfd_obs_z", "Z-score of observed relative FD",
      "rpd_obs_z", "Z-score of observed relative PD",
      "richness_obs_p_upper", "Richness percentile",
      "fd_obs_p_upper", "FD percentile",
      "pd_obs_p_upper", "PD percentile",
      "pe_obs_p_upper", "PE percentile",
      "pd_signif", "Significance of PD compared to null distribution, two-sided test at alpha 0.05",
      "rpd_signif", "Significance of RPD compared to null distribution, two-sided test at alpha 0.05",
      "fd_signif", "Significance of FD compared to null distribution, two-sided test at alpha 0.05",
      "rfd_signif", "Significance of RFD compared to null distribution, two-sided test at alpha 0.05",
      "pe_signif", "Significance of PE compared to null distribution, two-sided test at alpha 0.05",
      "taxonomic_cluster", "Membership in cluster based on taxonomic distances",
      "phylo_cluster", "Membership in cluster based on phylogenetic distances",
      "endem_type", "Endemism type determined by CANAPE",
      "lat_area", "Area of 1-degree latitudinal band around grid-cell",
      "temp", "Annual mean temperature, in units of degrees celsius * 10",
      "temp_season", "Temperature seasonality",
      "precip", "Annual precipitation, measured in mm",
      "precip_season", "Precipitation seasonality",
      "percent_apo", "Percent apomictic taxa"
    ))
```

@@

MD5 checksum: `r tools::md5sum(here::here(japan_ferns_biodiv_dryad_file))`

@@

```{r comm-ferns-dryad-file}
comm_ferns_dryad <- read_csv(here::here(comm_ferns_dryad_file))
# Check that no data are missing
assert(comm_ferns_dryad, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(comm_ferns_dryad_file)` (contained in "results.zip"): Community
matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for
biodiversity analysis. Same as japan_ferns_comm_full.csv, but grid-cells with
redundancy <0.1 (indicating under-sampling) excluded.

Number of variables: `r ncol(comm_ferns_dryad)`

Number of cases/rows: `r nrow(comm_ferns_dryad)`

@@

MD5 checksum: `r tools::md5sum(here::here(comm_ferns_dryad_file))`

@@

```{r shape-ferns-dryad-file}
shape_ferns_dryad <- sf::st_read(here::here(shape_ferns_dryad_file))
# Check that no data are missing
assert(shape_ferns_dryad, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(shape_ferns_dryad_file)` (contained in "results.zip"): Location
of grid-cells (sites) for native, non-hybrid ferns in Japan used for
biodiversity analysis. Same as japan_ferns_shape_full.gpkg, but grid-cells with
redundancy <0.1 (indicating under-sampling) excluded.

Number of variables: `r ncol(shape_ferns_dryad)`

Number of cases/rows: `r nrow(shape_ferns_dryad)`

@@

```{r shape-ferns-dryad-file-cols, results = "asis"}
  shape_ferns_dryad %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

MD5 checksum: `r tools::md5sum(here::here(shape_ferns_dryad_file))`

@@

```{r fern-traits-dryad-file}
fern_traits_dryad <- read_csv(here::here(fern_traits_dryad_file))
```

`r fs::path_file(fern_traits_dryad_file)` (contained in "results.zip"): Trait
matrix of native, non-hybrid ferns in Japan used for functional biodiversity
analysis.

Number of variables: `r ncol(fern_traits_dryad)`

Number of cases/rows: `r nrow(fern_traits_dryad)`

@@

MD5 checksum: `r tools::md5sum(here::here(fern_traits_dryad_file))`

@@

```{r japan-fern-phylogram-dryad-file}
japan_fern_phylogram_dryad <- ape::read.tree(here::here(japan_fern_phylogram_dryad_file))
```

`r fs::path_file(japan_fern_phylogram_dryad_file)` (contained in
"results.zip"): Maximim-likelihood phylogeny of native, non-hybrid ferns in
Japan in newick format inferred with maximum likelihood IQ-TREE v1.6.12
[@Nguyen2015]. Values at nodes are bootstrap support values calculated using
1000 replicates of ultrafast bootstrap [@Nguyen2015].

Number of tips: `r ape::Ntip(japan_fern_phylogram_dryad)`

@@

MD5 checksum: `r tools::md5sum(here::here(japan_fern_phylogram_dryad_file))`

@@

```{r japan-fern-dryad-uncollapsed-file}
japan_fern_tree_uncollapsed_dryad <- ape::read.tree(here::here(japan_fern_tree_uncollapsed_dryad_file))
```

`r fs::path_file(japan_fern_tree_uncollapsed_dryad_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015]. Clades consisting of identical OTUs not collapsed.

Number of tips: `r ape::Ntip(japan_fern_tree_uncollapsed_dryad)`

@@

MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_uncollapsed_dryad_file))`

@@

```{r japan-fern-dryad-file}
japan_fern_tree_dryad <- ape::read.tree(here::here(japan_fern_tree_dryad_file))
```

`r fs::path_file(japan_fern_tree_dryad_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015]. Clades consisting of identical sequences have been
collapsed to polytomies with zero branch length between OTUs.

Number of tips: `r ape::Ntip(japan_fern_tree_dryad)`

@@

MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_dryad_file))`

>>>>

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ebihara_2019_zip_file",
    "ftol_zip_file",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_figshare_file",
    "comm_ferns_figshare_file",
    "shape_ferns_figshare_file",
    "fern_traits_figshare_file",
    "japan_fern_phylogram_figshare_file",
    "japan_fern_tree_figshare_file",
    "japan_fern_tree_uncollapsed_figshare_file",
    "refs_yaml"
    )
  )
```

@@

```{r japan-ferns-biodiv-figshare}
japan_ferns_biodiv_figshare <- read_csv(here::here(japan_ferns_biodiv_figshare_file))
```

`r fs::path_file(japan_ferns_biodiv_figshare_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. 
Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Includes one row with missing
environmental data and one outlier for % apomixis that were removed prior to
spatial modeling analysis in Nitta et al. 2021.

Number of variables: `r ncol(japan_ferns_biodiv_figshare)`

Number of cases/rows: `r nrow(japan_ferns_biodiv_figshare)`

@@

```{r japan-ferns-biodiv-figshare-cols, results = "asis"}
  japan_ferns_biodiv_figshare %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "lat", "Latitude of grid-cell centroid",
      "long", "Longitude of grid-cell centroid",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "fd_obs_z", "Z-score (standard effect size) of observed functional diversity (FD)",
      "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
      "pe_obs_z", "Z-score of observed phylogenetic endemisim (PE)",
      "rfd_obs_z", "Z-score of observed relative FD",
      "rpd_obs_z", "Z-score of observed relative PD",
      "richness_obs_p_upper", "Richness percentile",
      "fd_obs_p_upper", "FD percentile",
      "pd_obs_p_upper", "PD percentile",
      "pe_obs_p_upper", "PE percentile",
      "pd_signif", "Significance of PD compared to null distribution, two-sided test at alpha 0.05",
      "rpd_signif", "Significance of RPD compared to null distribution, two-sided test at alpha 0.05",
      "fd_signif", "Significance of FD compared to null distribution, two-sided test at alpha 0.05",
      "rfd_signif", "Significance of RFD compared to null distribution, two-sided test at alpha 0.05",
      "pe_signif", "Significance of PE compared to null distribution, two-sided test at alpha 0.05",
      "taxonomic_cluster", "Membership in cluster based on taxonomic distances",
      "phylo_cluster", "Membership in cluster based on phylogenetic distances",
      "endem_type", "Endemism type determined by CANAPE",
      "lat_area", "Area of 1-degree latitudinal band around grid-cell",
      "temp", "Annual mean temperature, in units of degrees celsius * 10",
      "temp_season", "Temperature seasonality",
      "precip", "Annual precipitation, measured in mm",
      "precip_season", "Precipitation seasonality",
      "percent_apo", "Percent apomictic taxa"
    ))
```

@@

MD5 checksum: `r tools::md5sum(here::here(japan_ferns_biodiv_figshare_file))`

@@

```{r comm-ferns-figshare-file}
comm_ferns_figshare <- read_csv(here::here(comm_ferns_figshare_file))
# Check that no data are missing
assert(comm_ferns_figshare, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(comm_ferns_figshare_file)` (contained in "results.zip"): Community
matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for
biodiversity analysis. Same as japan_ferns_comm_full.csv, but grid-cells with
redundancy <0.1 (indicating under-sampling) excluded.

Number of variables: `r ncol(comm_ferns_figshare)`

Number of cases/rows: `r nrow(comm_ferns_figshare)`

@@

MD5 checksum: `r tools::md5sum(here::here(comm_ferns_figshare_file))`

@@

```{r shape-ferns-figshare-file}
shape_ferns_figshare <- sf::st_read(here::here(shape_ferns_figshare_file))
# Check that no data are missing
assert(shape_ferns_figshare, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(shape_ferns_figshare_file)` (contained in "results.zip"): Location
of grid-cells (sites) for native, non-hybrid ferns in Japan used for
biodiversity analysis. Same as japan_ferns_shape_full.gpkg, but grid-cells with
redundancy <0.1 (indicating under-sampling) excluded.

Number of variables: `r ncol(shape_ferns_figshare)`

Number of cases/rows: `r nrow(shape_ferns_figshare)`

@@

```{r shape-ferns-figshare-file-cols, results = "asis"}
  shape_ferns_figshare %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

MD5 checksum: `r tools::md5sum(here::here(shape_ferns_figshare_file))`

@@

```{r fern-traits-figshare-file}
fern_traits_figshare <- read_csv(here::here(fern_traits_figshare_file))
```

`r fs::path_file(fern_traits_figshare_file)` (contained in "results.zip"): Trait
matrix of native, non-hybrid ferns in Japan used for functional biodiversity
analysis.

Number of variables: `r ncol(fern_traits_figshare)`

Number of cases/rows: `r nrow(fern_traits_figshare)`

@@

MD5 checksum: `r tools::md5sum(here::here(fern_traits_figshare_file))`

@@

```{r japan-fern-phylogram-figshare-file}
japan_fern_phylogram_figshare <- ape::read.tree(here::here(japan_fern_phylogram_figshare_file))
```

`r fs::path_file(japan_fern_phylogram_figshare_file)` (contained in
"results.zip"): Maximim-likelihood phylogeny of native, non-hybrid ferns in
Japan in newick format inferred with maximum likelihood IQ-TREE v1.6.12
[@Nguyen2015]. Values at nodes are bootstrap support values calculated using
1000 replicates of ultrafast bootstrap [@Nguyen2015].

Number of tips: `r ape::Ntip(japan_fern_phylogram_figshare)`

@@

MD5 checksum: `r tools::md5sum(here::here(japan_fern_phylogram_figshare_file))`

@@

```{r japan-fern-figshare-uncollapsed-file}
japan_fern_tree_uncollapsed_figshare <- ape::read.tree(here::here(japan_fern_tree_uncollapsed_figshare_file))
```

`r fs::path_file(japan_fern_tree_uncollapsed_figshare_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015]. Clades consisting of identical OTUs not collapsed.

Number of tips: `r ape::Ntip(japan_fern_tree_uncollapsed_figshare)`

@@

MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_uncollapsed_figshare_file))`

@@

```{r japan-fern-figshare-file}
japan_fern_tree_figshare <- ape::read.tree(here::here(japan_fern_tree_figshare_file))
```

`r fs::path_file(japan_fern_tree_figshare_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015]. Clades consisting of identical sequences have been
collapsed to polytomies with zero branch length between OTUs.

Number of tips: `r ape::Ntip(japan_fern_tree_figshare)`

@@

MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_figshare_file))`
